---
title: "Lab 3"
author: "Khoa Tran"
date: "01/28/20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(oilabs)
```

* * *

### Exercise 1

```{r}
coin_outcomes <- c("heads", "tails")
set.seed(3111)
sim_fair_coin <- sample(coin_outcomes, size = 100, replace = TRUE)
sim_fair_coin %>% table
```

Of these 100 flips, 48 came up as heads.

### Exercise 2

```{r}
set.seed(111)
sim_unfair_coin <- sample(coin_outcomes, size = 100, replace = TRUE, 
                          prob = c(0.2, 0.8))
sim_unfair_coin %>% table
```

### Exercise 3

```{r}
set.seed(555)
dice_outcomes <- c(1,2,3,4,5,6)
sim_fair_die <- sample(dice_outcomes, size = 20, replace = TRUE)
sim_fair_die %>% table
```

### Exercise 4

```{r}
kobe_basket <- read.delim("http://anna-neufeld.github.io/Stat311/oiLabs/Week4/kobe_basket.csv", header=TRUE, sep=",")
calc_streak <- function(shots) {
  hits <- c(0, shots=="H", 0)
  misses <- which(hits==0)
  streaklengths <- diff(misses)-1
  return(data.frame(length= streaklengths))
}
```

```{r}
kobe_streak <- calc_streak(kobe_basket$shot)
ggplot(data = kobe_streak, aes(x = length)) +
  geom_histogram(binwidth=1)
kobe_streak %>% summarize(mean = mean(length), median = median(length))
```

The median is 0 and is a more accurate measurement for the measure of the center of the data because it is a right skew, making the mean more susceptible to the outliers and can't be an accurate measure of the center.

### Exercise 5

    
```{r}
shot_outcomes <- c("H", "M")
sim_basket <- sample(shot_outcomes, size = 100, replace = TRUE, prob = c(0.45, 0.55))
sim_basket %>% table
```
    

### Exercise 6
```{r}
set.seed(500)
shot_outcomes <- c("H", "M")
sim_basket <- sample(shot_outcomes, size = 133, replace = TRUE, prob = c(0.45, 0.55))
sim_basket %>% table
```

### Exercise 7

Of the 133 shots, 61 shots were hits.

### Exercise 8
```{r}
sim_streak <- calc_streak(sim_basket)
max(kobe_streak)
```

### Exercise 9

```{r}
max(sim_streak)
```

### Exercise 10
```{r}
ggplot(data=sim_streak, aes(x=length)) + geom_histogram(binwidth=1)
ggplot(data=kobe_streak, aes(x=length)) + geom_histogram(binwidth=1)
```

For both models. the distribution was very similar as an independent shooter with 0.45 probability shots made can just easily have a "hot hand". Using this comparision, there is no substainial evidence of Kobe having a "hot hand".
