---
title: "The Normal Distribution"
author: "Khoa Tran"
date: "02/06/2020"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(oilabs)
fastfood <- read.csv("http://anna-neufeld.github.io/Stat311/oiLabs/Week5/fastfood.csv")
mcdonalds <- fastfood %>%
  filter(restaurant == "Mcdonalds")
dairy_queen <- fastfood %>%
  filter(restaurant == "Dairy Queen")
```

* * *

### Exercise 1: 

```{r}
ggplot(data = mcdonalds, aes(x = cal_fat)) + geom_histogram(binwidth = 80)
ggplot(data = dairy_queen, aes(x = cal_fat)) + geom_histogram(binwidth = 80)
```

### Exercise 2:

The dairy queen distribution is more symmetric.

### Exercise 3: 

```{r}
dqmean <-mean(dairy_queen$cal_fat)
dqsd   <- sd(dairy_queen$cal_fat)
ggplot(data = dairy_queen, aes(x = cal_fat)) +
        geom_blank() +
        geom_histogram(binwidth = 80, aes(y = ..density..)) +
        stat_function(fun = dnorm, args = c(mean = dqmean, sd = dqsd), col = "tomato")
```

According to this plot, it does appear that the data follows a nearly normal distribution.

### Exercise 4

```{r}
ggplot(data = dairy_queen, aes(sample = cal_fat)) + 
  stat_qq() + 
  stat_qq_line()
sim_norm <- rnorm(n = nrow(dairy_queen), mean = dqmean, sd = dqsd)
ggplot(data = NULL, aes(sample = sim_norm)) + 
  stat_qq() + 
  stat_qq_line()
```

All the points do not fall on the line for either graph. Both have some randomness to them.

### Exercise 5

```{r}
source("https://anna-neufeld.github.io/Stat311/oiLabs/Week5/qqnormsim.R")
qqnormsim(sample = cal_fat, data = dairy_queen)
```

They are nearly normal because the data from known normal distributions have relativey the same range of data points as the unknown distribution type data.

### Exercise 6

```{r}
source("https://anna-neufeld.github.io/Stat311/oiLabs/Week5/qqnormsim.R")
qqnormsim(sample = cal_fat, data = mcdonalds)
```

### Exercise 7

The normal Q-Q plot does not look similar to the ones created with simulated data. They don't provide evidence that the calories from fat at McDonalds are nearly normal.

### Exercise 8

```{r}
(nrow(dairy_queen %>% filter(cal_fat < 500))) / (nrow(dairy_queen))
```


### Exercise 9

```{r}
1 - pnorm(q = 500, mean = dqmean, sd = dqsd)
```

### Exercise 10

```{r}
x <- pnorm(q = 600, mean = dqmean, sd = dqsd)
y <-pnorm(q = 400, mean = dqmean, sd = dqsd)
x-y
```


### Exercise 11

```{r}
mcmean <-mean(mcdonalds$cal_fat)
mcsd   <- sd(mcdonalds$cal_fat)
theo <- (1 - pnorm(q = 900, mean = mcmean, sd = mcsd))
empir <- (nrow(mcdonalds %>% filter(cal_fat >900))) / (nrow(mcdonalds))
print (theo)
print(empir)
```

The theoretical and empirical probability differs by a factor of 10.





