---
title: "Stat 311, HW6"
author: "Your Name"
date: "Wednesday Mar 4"
output: oilabs::lab_report
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(oilabs)
```

### Instructions
+ This homework is due on Wednesday Mar 4 by midnight. No late work is excused unless prior arrangements have been made with the instructor. The dropbox will remain open for 8 hours past the deadline. However, submissions made during this time will be penalized by at least 20% and maybe even as much as 50%.
+ Please answer ALL questions (including those that do not involve R) in the locations marked in this template. Remember to periodically **knit** your document to check that the output appears as you want it to; you will be turning in your knit html document. 
+ You will need to upload this document in CANVAS. Please be sure to check that your file was uploaded correctly. It does not hurt to screenshot verification of the upload as CANVAS can be glitchy on occasion.
+  Please answer the questions in the order in which they are posed. Write in complete sentences, and support your answers where asked. 

### Reading  (NOT OPTIONAL)
+ Sections 5.2, 5.3, 6.1 in OpenIntro (4th Addition) 

* * *
### Exercise 1: True or false and explain briefly

a) A type 1 error occurs anytime we reject the null hypothesis

  False, a type 1 error occurs when one rejects the null hypothesis in favor of the alternative hypothesis when the null hypothesis is true. A type 1 error does not occur anytime, rather in the specific instance I specified.

b) If the p-value is 0.03, the corresponding test will reject at a level of significance of 0.02.

  False. Since the p-value is larger than the level of significance, the data is not surprising enough to reject the null hypothesis. P-value is a test that implements the lowest significance level to reject the null hypothesis, so p-value >= level of significance does not reject.

c) The p-value is a random variable. 

  True, the p-value is a random variable. The definition of a random variable is a variable whose values depend on outcomes of a random phenomenon. The p-value fits that definition because it is based on the distribution of a test statistic that is supposed to be randomly sampled. Because the test statistic is assumed to have been randomly sampled and the p-value is a variable whose outcome depends on that randomness, it is a random variable. 

d) We can interpret the confidence interval (0.393, 0.553) by saying there is a 95\% chance that the population \% will fall between 0.393 and 0.553.
 
  False, we can interpret the confidence interval in that we are 95% percent confident that the true proportion will lie between .393 and .553, which is a concise difference. 

### Exercise 2: Pew Research

In a Pew Research Center poll of 745 randomly selected American adults, 589 said that it is morally wrong to not report all income on tax returns. 
  
a) In words, what is the parameter of interest? 
 
  The parameter of interest is the proportion of American adults who say that it is morally wrong to not report all income on tax returns. 

b) Use sample results to construct a 98\% confidence interval estimate of the parameter. 

  p_hat is 589/745, which is approximately .7906.
  To find the standard error, need to take (1-p_hat)*(p_hat), divide by n, and take square root. 
  n is 745, and SE is approximately .015. Now, need to multiply standard error by the confidence level. The confidence level multiplier for a 98% confidence interval is 2.326, so multiply SE by 2.326 to get approximately .0347.The interval is p_hat +/- .0347, so the confidence interval is (.7559, .8253). 
  The confidence statement is as follows: We are 98% confident that the proportion of American adults who say that is morally wrong to not report all income tax returns falls in the interval of .7559 to .8253.

c) Assume that we plan to conduct another poll to determine whether the given sample results are accurate. How many adults would we need to poll in order to have 98\% confidence that the (quick) margin of error is 0.02? Show your calculations below.

  Around 2,240 adults. To do that, it's necessary that I reverse the work I did to find the margin of error. The margin of error equation is (1-p_hat)*(p_hat), divide by n,take square root, and multiply by confidence level multiplier to find margin of error. Now we have margin of error to be .02. Divide .02 by the confidence level multiplier for 98% which 2.326. Then square it to get rid of the square root. Multiply both sides by n to get n on the left side, and then divide both sides by (margin of error/cl multiplier)^2 to isolate n on the left. The answer for n turns out to be (p_hat)(1-p_hat)(cl multiplier^2) all over (margin of error)^2, which comes out to be approximately 2,239.2.
 
d) Unfortunately, an important condition necessary for inference may be violated here. what might this be?   (No help will be given on this question. Please do not discuss this with anyone.)

  
There are three important conditions to check in order to infer: the random condition, the normal condition, and the independence condition. The random condition requires that the sample be done randomly, which in this case is done. The normal condition requires that np >= 10 and n(1-p) >= 10, which is also fulfilled. Last left is the independence condition, which is if you sample without replacement, one must sample 10% or less. The American adult population is atleast 209 million, so 2,400 is definitely less than 10%. All these conditions seem like they stay intact, but the condition that seems to be most violated is the independence condition, due to the fact that the sampling is done without replacement.


### Exercise 3: Type 1 and Type 2 errors

a) J \& T each want to test if the proportion of adults in their neighborhood who have graduated from high school is 0.94, as claimed in the newspaper. J takes a random sample of 200 adults and uses $\alpha=0.05$. T takes a random sample of 500 adults and uses $\alpha=0.05$. Suppose the newspaper's percentage is actually right.

(i) Is it possible for either J or T to make a type 1 error? If yes, who is more likely to do so? If no, why not?

Yes, it is possible for either to make a type 1 error, since we know the null hypothesis is true, it is possible that one of the two or both rejects the null hypothesis even though it is actually right. Out of the two, both are equally likely to commit a type 1 error, and that's due the null being true. When the null is true, p-values are uniformly distributed between 0 and 1, so the probability that the null is rejected is alpha. Both of them use an alpha of .05, and that does not depend on sample size. So both are equally likely.

(ii) Is it possible for either J or T to make a type 2 error? If yes, who is more likely to do so? If no, why not?

It is not possible for J or T to make a type 2 error, because in order to make a type 2 error, the alternative hypothesis should be actually right. We know that the null hypothesis, which is the newspaper's percentage is actually right, so a type 2 error cannot occur.

b) J \& T each want to test if the proportion of adults in their neighborhood who took chemistry in high school is 0.25, as claimed in the newspaper. J takes a random sample of 200 adults and uses $\alpha=0.05$. T takes a random sample of 500 adults and uses $\alpha=0.05$. Suppose the newspaper's percentage is actually wrong.

(i) Is it possible for either J or T to make a type 1 error? If yes, who is more likely to do so? If no, why not?

It is not possible for J or T to make a type 1 error, because in order to make a type 1 error, the null hypothesis should be actually right. We know that the null hypothesis, which is the newspaper's percentage is actually wrong, so a type 1 error cannot occur.

(ii) Is it possible for either J or T to make a type 2 error? If yes, who is more likely to do so? If no, why not?

It's possible for both to make a type 2 error, and that's due to the fact that we know the null hypothesis is wrong, which is .25. It is possible that one or both may fail to reject the null hypothesis. As for who's more likely to do so, I again believe J is more likely to make a type 2 error since we know the expected value isn't the newspaper number, we know thr null hypothesis is wrong. This would mean p has to be below .05, which has statistical significance in that observations made would be considered extreme if the null were true. The observations being made should be classified as extreme because the null hypothesis is wrong. By having a large sample size, it becomes more likely that the resu;ts' average out to the actual right proportion, i.e. the expected value. That's according to the Law of Large Numbers. By having a larger sample size, it becomes more likely that the p is less than .05 because the results are closer to the actual proportion, which makes it more likely that one rejects the null hypothesis. Due to that, since J has less of a sample size, she's more likely to err.


### Exercise 4: Working backwards

Problem 5.24 on page 203

To find the sample proportion for the p-value equaling .01 with a sample size of 1,429, need to find the z-score corresponding to the p-value of .01. The z-score that corresponds with .01 for a two-tailed test is approximately 2.576. Now we have the Z-score, we can find the sample proportion by using the Z-score equation. The Z-score equation is sample proportion - mean proportion(which is the null hypothesis proportion of .9 in this case) divided by standard error, which is (1-mean proportion)*(mean proportion) divided by sample size. After solving for sample proportion, the answer is .9204.


### Exercise 5: What happens when the null is false?

For this exercise, we are working with the same dataset and setup as Lab 6. You should load the data and complete the set-up steps.

```{r}
download.file("http://www.openintro.org/stat/data/nc.RData", destfile = "nc.RData")
load("nc.RData")
nc_smoker <- nc %>% filter(habit=="smoker")
phat <- mean(nc_smoker$lowbirthweight=="low")
```

Recall that on lab 3, we were working with a random sample of 126 smoking mothers where 14.3% had low-weight babies ($\hat{p}=0.143$). We tested the null hypothesis that $p=0.1$ and failed to reject the null using our sample. To better understand the meaning of a p-value, we then simulated 100,000 samples from a hypothetical population in a world where the null hypothesis was true. 

Now we will imagine we live in a world where the null hypothesis is false. Suppose 
that $p=0.13$, meaning that smoking mothers really are more likely to have underweight babies than the population as a whole. Let's create a hypothetical population:  

```{r}
popsize <- 100000
new_pop <- data.frame(lowbirthweight = c(rep(0, 0.87*popsize), rep(1, 0.13*popsize))) 
```

Now let's repeat our previous simulation. Note that we as researchers don't **know** that the truth is 0.13, so this information will not be plugged in anywhere. We are still testing the null that p=0.1.

```{r}
library(infer)
set.seed(311)
p0 <- 0.1
n <- 126
true_sd <- sqrt(p0*(1-p0)/n)
new_results <- new_pop %>%
        infer::rep_sample_n(size = 126, reps = 10000) %>%
        summarise(phat = mean(lowbirthweight),
                  se = sqrt(phat*(1-phat)/126),
                  me = 1.96 * se,
                  lower = phat - me,
                  upper = phat + me,
                  CI_covers_truth = (lower <= 0.13 && upper >= 0.13),
                  CI_covers_null = (lower <= p0 && upper >= p0),
                  zscore = (phat-p0)/se,
                  pval = 2*(1-pnorm(abs(zscore))),
                  reject_null = pval <= 0.05)
```

a. What was the largest observed $\hat{p}$? The smallest?

```{r}
min(new_results$phat)
max(new_results$phat)
```
The largest observed phat was approximately .254, while the smallest observed phat was approximately .032.

b. How many times did you make a type 1 error?

```{r}

```
0, because in order to make a type 1 error, the null hypothesis has to be true. In this question, we're living in a world where the null hypothesis is false, so it isn't possible to make a type 1 error.


c. How many times did you make a type 2 error?

```{r}
interval <- filter(new_results, reject_null == "FALSE")
nrow(interval)
```
The type 2 error was made 8,596 times.

d. Would it have been reasonable to observe a sample such as our original sample (``nc`` dataset) in a population where $p=0.13$? To explain your answer, make a histogram of all of the ``phats`` in ``new_results`` and add a vertical line representing your original sample.

```{r}
ggplot(data = NULL, aes(x=new_results$phat)) + geom_histogram(binwidth = 1/126) +  geom_vline(xintercept= phat, col = "red")
```
It would be reasonable to observe a sample such as the original one in a population where p=.13, because it's not too far from the mean in the normal distribution and falls well within two standard deviations away.


e. On the lab, you failed to reject the null that p=0.1. Based on this simulation, are you convinced that p really is 0.1? In other words, did you prove the null? Explain your answer, using your results from part d for support. 

I am not convinced that p really is .1, and that's because in order to prove the null, there is not enough evidence in the above graph where phat is around .14 and p is .1, the distribution shows .14 is not extreme. That means we cannot reject the null hypothesis. That does not mean however we proved the null. Rather we have proved the null hypothesis could be possibly right. There's too much uncertainty to say the null has been proved, when in fact we've now failed to reject p = .13 or p = .1, which means we haven't proven any nulls at all.




