---
title: "Stat 311, HW4"
author: "Ravi Sangani"
date: "Wednesday Feb 12"
output: oilabs::lab_report
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(oilabs)
```

### Instructions
+ This homework is due on Wednesday Feb 12 by midnight. No late work is excused unless prior arrangements have been made with the instructor. The dropbox will remain open for 8 hours past the deadline. However, submissions made during this time will be penalized by at least 20% and maybe even as much as 50%.
+ Please answer ALL questions (including those that do not involve R) in the locations marked in this template. Remember to periodically **knit** your document to check that the output appears as you want it to; you will be turning in your knit html document. 
+ You will need to upload this document in CANVAS. Please be sure to check that your file was uploaded correctly. It does not hurt to screenshot verification of the upload as CANVAS can be glitchy on occasion.
+  Please answer the questions in the order in which they are posed. Write in complete sentences, and support your answers where asked. 

### Reading  (NOT OPTIONAL)
+ Sections 3.4, 3.5, 4.1 in OpenIntro (4th Addition) 

* * *

### Exercise 1: Keno

Suppose you are going to Las Vegas to play Keno. Your favorite bet is $1 on a single number. When you win, they give you $3. When you lose, you get nothing. You have a 1 out 4 chance of winning. 

a. Build a probability model for $X=\text{net gain}$ in one play. (Net gain is the amount you receive minus the cost of playing, and yes, it can be -1)

   *State the values and the probabilities here*
   
   Outcome Values      Probability
   
   + $2                  .25
   - $1                  .75

b. What is the expected value of $X$? Interpret this number with a fully formed sentence.

   *Write expected value. Then write a separate sentence interpreting what this number tells you.*
   The expected value is E($X$) = (2 * .25) + (-1 * .75) = -.25. This means that you would expect to be 25 cents down after every play.
   

c. Is Keno a *fair* game. Why or why not?

   *Write whether it is fair or unfair. Then start a separate sentence giving your reasoning*
   Keno is an unfair game. This is because the cost to play isn't equal to the expected value. Rather, the cost to play is significantly higher than the expected value, 1 dollar vs 25 cents.

d. What is the standard deviation $X$? Interpret this number with a fully formed sentence. 

   *Write standard deviation. Then write a separate sentence interpreting what this number tells you.*
   SD($X$) = sqrt( ((2 - (-.25))^2 * .25) + ((-1 - (-.25))^2 * .75)) = 1.29
   This number tells you that roughly 68% of the data, regarding the net gain/loss after one play, will fall within -1.54 and +1.04.


# Exercise 2: More Fun Playing Keno

Despite your better judgment, you play Keno under the terms specified previously. Let $\bar{X}$ denote the average net gain after 100 plays. That is:
\[ \bar{X} = \frac{1}{100}(X_1 + X_2 + \dots + X_{100}) \]
where $X_i$ is the net gain from game $i$. 

a. $\bar{X}$ is a linear combination of the $X_i$'s. What are the coefficients of the linear combination?
 
   *Write coefficient here*
   1/100

b. Calculate the expected value of $\bar{X}$. State any results you are using.

   *Write expected value. Starting on a separate line, justify*
   The expected value for one play is -0.25. For the average net gain per play after 100 plays, the expected value would be also -0.25, because to find the value we essentially have to find the average of all the expected values. The expected value doesn't change for any of the 100 plays, so we are doing 100*(-0.25) / 100. 
   

c. Calculate the standard deviation of $\bar{X}$. State any results you are using. Are you making any assumptions to do this calculation? State those as well.

   *Write standard deviation. Starting on a separate line, justify*
   0.129
   
   The general equation for variance is E(X^2) - E^2(X). To carry this calculation out for $\bar{X}$ , we use the previous variance result for just $X$ in the calculation. With a linear combination like $\bar{X}$, the variance is a^2 * V($\bar{X}$). To evaluate the latter portion, I do a summation calculation of V($X$), which is already known, 100 times and then multiply this by 1/(100^2) which fulfills the a^2 portion. Then I take the square root to get the standard deviation. 
   

d. Suppose $\bar{X}$ is the average based on 1,000 (not 100) plays. How do the calculations from parts b and c change? 

   *Write one sentence stating how expected value and standard deviation change.*
   The expected value would remain the same because the number of times we do the summation is always inverse to the coefficient, when calculating E($X$). However, the standard deviation would be reduced as the variance is reduced by a factor of ten as the a^2 term would now be a factor of ten greater.
   

e. The law of large numbers states that averages of results from a large number of trials stabilizes around what is expected. Use what you have learned from part (d) to explain why this is so.

   *Write two sentences to explain*
   The average of results from a large number of trials would stabalize around what is expected as the standard deviation decreases with more trials. Therefore, there is less degree of variability with more trials. 


### Exercise 3: Think Inside the Box!

 Consider a box which contains some tickets. Suppose a fraction, $p$, of the tickets have a "1" written on them; the remaining have a "0".
 
 a. Let $X$ denote the number on a ticket which is randomly drawn from the box. Fill in the first blank in the following sentence with the expected value of $X$ and the second blank with the standard deviation of $X$.
 

    We expect X to be ___ p______ give or take ____$\sqrt{(p(1-p))}$________
 
   *Write a few sentences to explain how you calculated the numbers. You could include a table showing the probability distribution of X and then show how to calculate E(X) and SD(X) from the entries in the table*
   
   Outcome Value        Probability 
   1                    p
   0                    1-p

To calculate E(X), I did ((1 * p) + (0 * (1-p))) which turns out to be p. This is the probability multiplied the outcome value corresponding to it, summed up over all outcome values. To calculate the standard deviation, I took the square root of E(X^2) - E^2(X). The first term evaluates to p and the second to p^2. Therefore, the final value is sqrt(p(1-p))
   
 
 b. Let $\bar{X}$ denote the fraction of 1's in 10 draws made with replacement from the box. That is,
 \[ \bar{X} = \frac{X_1 + X_2 + \dots + X_{10}}{10} \]
 
    where $X_i$ denotes the number on the $ith$ draw. Fill in the first blank in the following sentence with the expected value of $\bar{X}$ and the second blank with the standard deviation of $\bar{X}$.  (To write the standard deviation symbolically, use the dollar sign like this:$\sqrt{ab/c}$ for instance)
 
    We expect $\bar{X}$ to be ___p______ give or take ____$\sqrt{((p(1-p))/10)}$________
 
   *Write a few sentences to explain how you calculated the numbers. You could include rules you used to arrive at your answers as well.*
   The expected value would still be p. With this situation, the number of times the summation happens for E($\bar{X}$) is inverse to the coefficient, and therefore the expected value will always be p. The standard deviation is the square root of the variance, which is equal to E(X^2) - E^2(X). However, because we are looking at $\bar{X}$, we can take the value for one draw's variance, and do the average for 10 times, essentially multiplying the value by 10, and then multiply by a^2, which is 1/(10^2) = 1/100. The final value is sqrt((p(1-p))/10). For the symbolic SD, a = p, b = 1-p, c = 10.

### Exercise 4. Density Curves

a) In each row of this table, assume the distribution is normal. Knowing any two of the mean, standard deviation, Q1 or Q3 is enough to determine the other two. Complete the table by filling in numbers for (i)--(iv). (You may use pnorm and qnorm in R and also use R as a calculator.)

| Mean | SD | $Q_1$| $Q_3$ |
|:--:  |:--:|:--:  |:--:   |
| 10   | 5  | 6.628| 13.37 |
| 150  | 30 | 120  |  180  |

```{r, echo=T}
qnorm(0.25, mean = 10, sd = 5, lower.tail = TRUE)
qnorm(0.75, mean = 10, sd = 5, lower.tail = TRUE)
x <- (120 + 180) / 2
x
std <- 150 - 120
std

```

b) A study conducted by the Kaiser Foundation compared blood pressures for women having different numbers of children. Below are sketches of the [density curves](https://drive.google.com/file/d/1RO9FuVA89-GlnPRAnQzz9_DZXn9attH4/view?usp=sharing) for blood pressure (measured in mm) in women with 2 or 4 children. 

(i) Is the *number of children* a woman has a numerical or categorical variable in the graphic. Support your answer.

*Write answer here*
The number of children is a categorical variable in the graphic as it seperates the women into different categories. Also, any calculations done on 2 and 4 don't have any meaningful results.

(ii) The caption for the y-axis is missing. What do the numbers on the $y$ axis tell us? (*Hint* look at the brief introduction to density histograms at the end of the slidedeck titled *Continuous Distributions*)

*Write answer here*
The numbers on the $y$ axis is the density of each bin. The density tells us the frequency of the x-value divided by the width of bin.

(iii) True or false and explain: the increase in the number of children has an *additive* effect on blood pressure. 

*Write true/false first and then support your answer*
True, the increase in the number of children has an additive effect on blood pressure. We can see in the graphic that with the distribution graph of women with 4 children, the continous distribution shifts to the right indicating that there is higher density of higher values of blood pressure. It becomes more more of a normal distribution as well. 


(iv) Roughly, what is the height of the density curve for women with 4 children? Show your work.

*Write answer first and then support with reason*

0.029. Area = 0.5 * base * height. We know the area is 0.5 for the right triangle from the right side of the normal distribution, so 0.5 = 0.5 * 35 * height. Height = 0.029.

(v) Blood pressures above 140 mmHg are considered unusually high. Which group has  a higher probability of exceeding 140-- the mothers of 2 children or 4 children? Explain how you know this. (You do not need to do calculations -- just use your eyeballs)

*Write which group and then support with reason*

The group of mothers with 4 children have a higher probability of exceeding 140 mmHg for blood pressure. I know this because an area spanning a certain width in a continuous distribution for a density curve corresponds to the probability. Visually, if we were to take the area under 140 for both curves, the graph for mothers with 4 children would yield a larger area; hence, a larger probability. 

### Exercise 5: Normal Approximations for Coin Flips

If we coss a toin $n$ times and let $X$ be the number of heads that we observe, then $X$ is called a *Binomial Random Variable*. Suppose that we have 50 fair coins, so $n = 50$ and $p$, the probability of seeing heads on each toss, is $0.5$. Our tosses are independent of one another. 

You know from your previous labs that the following code will simulate 1 realization of $X$. 
```{r}
coin_outcomes <- c(0,1)
X <- sum(sample(coin_outcomes, size=50, replace = TRUE))
```

To get an idea of the distribution of the random variable $X$, we want to repeat this experiment many times and observe many values of $X$. The following code records 1000 realizations of $X$.

```{r}
set.seed(111)
Xs <- replicate(1000, sum(sample(coin_outcomes, size=50, replace = TRUE)))
```

### Part A: 
Make a histogram of the realizations of $X$. Use a ``binwidth`` of 1. Note that since Xs is just a one-dimensional vector, not a dataframe, you should set ``data = NULL`` and ``aes(x=Xs)`` in your call to ``ggplot()``. Describe the shape of the distribution. Does it look approximately normal? 


The shape of the distribution is unimodal and symmetric. It looks approximately normal.

```{r}
ggplot(data=NULL, aes(x = Xs)) + geom_histogram(binwidth = 1)

```

### Part B:  
Make a density histogram of your data ``Xs`` and then overlay a density curve for the theoretical normal distribution with the same mean and standard deviation as the data ``Xs``. Use exercise 3 on lab 4 for help with a density histogram and density curve. 

```{r}
Xmean <-mean(Xs)
Xstd <- sd(Xs)
ggplot(data = NULL, aes(x = Xs)) +
        geom_blank() +
        geom_histogram(binwidth = 1, aes(y = ..density..)) +
        stat_function(fun = dnorm, args = c(mean = Xmean , sd = Xstd))
```

### Part C: 
What happens if the coin is not fair? Make a dataset called ``Xs_unfair`` that holds 1000 realizations of ``Number of heads in 50 tosses`` for a coin that only comes up heads (X=1) with probability 0.05. Use ``set.seed(333)``. (See exercise 2 on lab 3 for help simulating an unfair coin). 
```{r}
set.seed(333)
Xs_unfair <- replicate(1000, sum(sample(coin_outcomes, size=50, replace = TRUE, prob=c(0.95, 0.05))))

```

### Part D: 
Make a density histogram of your data ``Xs_unfair`` and then overlay a density curve for the theoretical normal distribution with the same mean and standard deviation as the data ``Xs_unfair``. What do you observe? Does the data look approximately normal? Use exercise 3 on lab 4 for help with a density histogram and density curve. 


The data is approximately normal because it closely follows the density curve of a theoretical normal distribution with same mean and standard deviation.
```{r}
meanx <-mean(Xs_unfair)
stdx <- sd(Xs_unfair)
ggplot(data = NULL, aes(x = Xs_unfair)) +
        geom_blank() +
        geom_histogram(binwidth = 1, aes(y = ..density..)) +
        stat_function(fun = dnorm, args = c(mean = meanx, sd = stdx))

```


