---
title: "Shootings Data Report"
author: "Khoa Tran"
date: "11/10/2019"
output: html_document
---
```{r setup, include = FALSE}
#Load the 'knitr' package
library(knitr)
#Load results from the analysis, without having the specific code
source("analysis.R")
```

## Summary Information

This is a summary of the mass shootings data in the US for 
2018 by location ([source](`r shootings_source`)). The total number of mass shootings
is `r total_num_shootings`. The total number of lives lost due to mass shootings is
`r total_lives_lost`. To analyze this data, I want to define the concept of "impacted" 
as the total number of people killed and injured. The city that was most impacted by mass shootings is
`r city_most_impacted`. The state that was the most impacted by mass shootings
is `r state_most_impacted`. The month with the most mass shootings is `r most_shootings_month`.

## Summary Table

This table shows how many people were impacted by shootings in each month,
with the total impacted people sorted in descending order.

```{r table, echo = FALSE}
#Display the summary table with the specified column names
kable(summarize_shootings_data, col.names = c("Month", "Total Affected"),
      align = "l")
```

#### Two Major Insights:
- The month December had the least people impacted by shootings
- The month June had the most people impacted by shootings

## Description of a Particular Incident

Out of all the shootings 2018, only one occurred in the city of
`location_of_incident`. This shooting occurred on the `r date_of_incident` 
at the geolocation (`r geolocation_of_incident`). 
There were `r num_injured_of_incident` people injured and `r num_killed_of_incident`
people killed during this incident. For more information about this shooting,
you can visit this [link](`r incident_extresource`).

## Interactive Map

This map uses circles of in a variety of sizes to show the relativity 
how many people were impacted by shootings during the year of 2018 by location. 
This map also allows for visuals of area had more or less shootings.

```{r interactice_map, echo=FALSE}
#Shows the interactive map
map_of_shootings
```

#### Two Major Insights:
- South Florida and New York had the most amount of people
impacted by shootings making it dangerous regions
- The east coast has significantly more shootings than the west coast and central US

## Fifteen Worst Shootings

#### Proposed Questions:
What was the fifteen biggest mass shootings in 2018?
Where were they located and how many people were impacted?

```{r twenty_biggest_shootings, echo = FALSE}
#Creates a horizontal bar chart
ggplot(fifteen_biggest_shootings) +
  geom_col(mapping = aes(x = location, y = impacted_num)) +
  labs( 
    #Add labels to x and y axis
    #Plot title
    title = "Fifteen Biggest Mass Shootings in 2018", 
    #Vertical X axis title
    x = "Location",
    #Horizontal Y axis title
    y = "Number of People Impacted" 
  ) +
  coord_flip() #Make horizontal bars with cities on the side for better visibility
```

#### Two Major Insights:
- Florida and Califonira have the most number of biggest mass shootings in 2018
- Parkland, Florida was the biggest and worst shooting in 2018 with more than 30 people impacted
